---
title: "recipes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Tools

* rjson

```{r echo = FALSE, message=FALSE, reformat=TRUE}
# install.packages("rjson")
require(data.table)
require(jsonlite)
require(dplyr)
require(tidyr)

```

## Recipes Analysis
Looking at data on 40k recipes, using ingredients + labeled cuisines to build a predictive classification model.  Testing model on 10k recipes unlabeled data set.

### Load & Format Data

* **Training Data**: 39774 recipes
* **Test Data**: 9944 recipes

```{r echo = FALSE, message=FALSE, reformat=TRUE}
## load training and testing data
train_file <- "data/train.json"
test_file <- "data/test.json"
train <- fromJSON(train_file)
test <- fromJSON(test_file)

train <- as_tibble(train)
test <- as_tibble(test)

#str(train)
#str(test)
```

Flattened Training Data - 1 row per ingredient
```{r echo = FALSE}
train_ingredients <- train %>%
  unnest(ingredients)

train_ingredients
```
Flattened Test Data - 1 row per ingredient
```{r echo = FALSE}

test_ingredients <- test %>%
  unnest(ingredients)

test_ingredients
```



### Summarize Data
```{r echo = FALSE}
summary(train_ingredients)
summary(test_ingredients)

```

#### How many total recipes are there?
```{r echo = FALSE}

train_recipe_count <- train_ingredients %>%
  summarise(n_recipes = n_distinct(id))

as.integer(train_recipe_count)

```

#### What are the different cuisines?  How many recipes are there in the training data?
```{r echo = FALSE}

train_cuisine_recipe_counts <- train_ingredients %>%
  group_by(cuisine) %>%
  summarise(n_recipes = n_distinct(id)) %>%
  arrange(desc(n_recipes))

train_cuisine_recipe_counts

```


#### What are the top ingredients per cuisine? How often do they appear in the training data?
```{r echo = FALSE}
train_ingredient_counts_by_cuisine <- train_ingredients %>%
  group_by(cuisine) %>%
  count(cuisine,ingredients) %>%
  arrange(cuisine, desc(n))

train_ingredient_frequencies_by_cuisine <- 
  merge(train_ingredient_counts_by_cuisine, train_cuisine_recipe_counts) %>%
  mutate(pct_of_cuisine_recipes = n/n_recipes)

train_ingredient_frequencies_by_cuisine %>%
  group_by(cuisine) %>%
  arrange(cuisine, desc(n)) %>%
   top_n(10)

```

#### What are the most distinctive ingredients per cuisine?
```{r echo = FALSE}
train_ingredient_counts <- train_ingredients %>%
  group_by(ingredients) %>%
  count(ingredients) %>%
  arrange(desc(n))

train_ingredient_frequencies <- 
  train_ingredient_counts %>%
  mutate(n_total_recipes = as.integer(train_recipe_count)) %>%
  mutate(pct_of_total_recipes = n/n_total_recipes)


merge(train_ingredient_frequencies_by_cuisine, train_ingredient_frequencies, by="ingredients") %>%
  mutate(pct_diff = round((pct_of_cuisine_recipes - pct_of_total_recipes),2)) %>%
  mutate(pct_of_cuisine_recipes = round(pct_of_cuisine_recipes,2)) %>%
  mutate(pct_of_total_recipes = round(pct_of_total_recipes,2)) %>%
  arrange(desc(pct_diff)) %>%
  # arrange(pct_diff) %>%
  select(ingredients, cuisine, pct_of_cuisine_recipes, pct_of_total_recipes, pct_diff) %>%
  top_n(20)

```

