---
title: "Titanic Analysis Learnings Summary"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Review material from Data Analysis and highlight key learnings


## (1) Initial Data Analysis
### Load data into R 
using `read.csv` method
```{r results = "hide"}
test <- read.csv("test.csv",header = TRUE)
train <- read.csv("train.csv",header = TRUE)
```

### Understand data

#### Look at Data Dictionary  
https://www.kaggle.com/c/titanic/data

Variable    | Definition           	 | Key                |
------------|------------------------|--------------------|
survival    | Survival               | 0 = No, 1 = Yes    |
pclass      |	Ticket class            |	1 = 1st, 2 = 2nd, 3 = 3rd |
sex         |	Sex	                    |                         |
age	| Age in years	| |
sibsp	| # of siblings / spouses aboard the Titanic	| |
parch	| # of parents / children aboard the Titanic	| |
ticket  |	Ticket number	  |
fare  |	Passenger fare	 |
cabin	| Cabin number	 |
embarked  |	Port of Embarkation	| C = Cherbourg, Q = Queenstown, S = Southampton  |
##### Variable Notes
**pclass**: A proxy for socio-economic status (SES)
1st = Upper
2nd = Middle
3rd = Lower

**age**: Age is fractional if less than 1. If the age is estimated, is it in the form of xx.5

**sibsp**: The dataset defines family relations in this way...
Sibling = brother, sister, stepbrother, stepsister
Spouse = husband, wife (mistresses and fiancés were ignored)

**parch**: The dataset defines family relations in this way...
Parent = mother, father
Child = daughter, son, stepdaughter, stepson
Some children travelled only with a nanny, therefore parch=0 for them.

##### Notes:
* `survival` is the classification variable


#### Look at data itself
```{r}
head(train)
```
##### Notes:
* why don't the column names match the data dictionary.  fix that.
```{r results = "hide"}
colnames(train) <- tolower(colnames(train))
colnames(test) <- tolower(colnames(test))
```


#### Look at data structure
```{r}
str(train)
```
We can see the data types.


#### Look at summary statistics.  Understand prevelance of missing / null data.

```{r}
summary(train)
```

#### Look at train vs. test splits. 
* About 2/3 train (1:891), 1/3 test (892:1309) 
* Union together using `rbind()` into a combined dataframe in which we store any newly engineered features so that they are available both when training and testing our models.   
* At this point we can convert data types if it makes analysis easier.  or make a new column with the data with new data type.
```{r results="hide"}
## combine test and train datasets
test.survived <- data.frame(survived = rep("None",nrow(test)),test[,])
combined <- rbind(train, test.survived)

## change datatypes
combined$pclass <- as.factor(combined$pclass)
combined$survived <- as.factor(combined$survived)
combined$name <- as.character(combined$name)
```


### explore relationships between data fields vs. classification field `survived`

**Goal**: make hypotheses for which features have predictive power and why

#### Load Tools:
* `gmodels`: use `CrossTable` to easily make tables of counts & proportions
* `dplyr`: for aggregating data like sql/pivot table for proportions
* `ggplot2`: use for plots
* `cowplot`: use for plotting multiple graphs side-by-side
* `stringr`: use for manipulating strings
```{r message=FALSE} 
library(dplyr)
library(gmodels)
library(ggplot2)
library(cowplot)
library(stringr)
```


#### survived

* training set: 
    + 549 (62%) perished
    + 342 (38%) survived
```{r results="hide"}
## using gmodels CrossTable()
CrossTable(train$survived)
```
```{r}
## using dplyr
combined[1:891,] %>% 
  group_by(survived) %>% 
  summarize(count=n()) %>% 
  mutate(pct = count/sum(count))
```

#### pclass
Predictive? **Yes**

* survival rates 
    + 63% in 1st class
    + 48% in 2nd class
    + 24% in 3rd class
```{r echo=FALSE, eval=FALSE}
combined[1:891,] %>% 
  group_by(pclass, survived) %>% 
  summarize(count=n()) %>% 
  mutate(pct = count/sum(count))
```
* % of passengers 
    + 3rd class: 55%
    + 1st: 24%
    + 2nd: 21%
```{r echo=FALSE, eval=FALSE}
combined[1:891,] %>% 
  group_by(pclass) %>% 
  summarize(count=n()) %>% 
  mutate(pct = count/sum(count), pct2 = count/sum(sum(count)))
```

```{r echo=FALSE}
pclass.summary <- combined[1:891,] %>% 
  group_by(pclass, survived) %>% 
  summarize(count=n()) %>% 
  mutate(pct = count/sum(count))

a1 <- ggplot(pclass.summary, aes(x= pclass, y = pct, fill = survived)) +
  geom_bar(stat="identity", width = 0.7)

a2 <- ggplot(pclass.summary, aes(x= pclass, y = count, fill = survived)) +
  geom_bar(stat="identity", width = 0.7)

cowplot::plot_grid(a1,a2, ncol=1, labels = "AUTO")
```

#### sex
Predictive: **Yes**

* survival rates 
    + 74% for female
    + 19% for male
```{r echo=FALSE, eval=FALSE}
combined[1:891,] %>% 
  group_by(sex, survived) %>% 
  summarize(count=n()) %>% 
  mutate(pct = count/sum(count))
```

```{r echo=FALSE}
sex.summary <- combined[1:891,] %>% 
  group_by(sex, survived) %>% 
  summarize(count=n()) %>% 
  mutate(pct = count/sum(count))

b1 <- ggplot(sex.summary, aes(x= sex, y = pct, fill = survived)) +
  geom_bar(stat="identity", width = 0.7)

b2 <- ggplot(sex.summary, aes(x= sex, y = count, fill = survived)) +
  geom_bar(stat="identity", width = 0.7)

cowplot::plot_grid(b1,b2, ncol=1, labels = "AUTO")
```

#### age

Predictive? Maybe

* 177 (```r round(177/891*100,2)```%) rows have null `age` value, so will be hard to use without imputing data somehow
```{r}
summary(combined[1:891, "age"])
```

```{r echo=FALSE, warning=FALSE}
ggplot(combined[1:891,] ,aes(x = ifelse(is.na(age),105, age),fill = survived)) + 
  geom_bar(position = "fill") + 
  stat_bin(bins = 10)
```

##### Notes:
* add pct survival rate by age bin


#### name --> title

Predictive? **Yes**

##### Name Patterns
*Mr.*: `<lname>, <title>. <fname>`

*Mrs.*: `<lname>, <title>. <husband fname> <husband mname> (<fname> <mname>)`

*Ms.*: `<lname>, <title>. <fname>`


Extract last names and titles from `name` and add to combined data set

```{r echo=FALSE} 

## returns a list of names, where each name is separated into a vector with the part before and after the comma
name.splits <- str_split(combined[,"name"],', ')
#str_split(combined[1:5,"name"],', ')[[2]][2]

## pull first part of each name before comma (last name)
last.names <- sapply(name.splits, '[', 1)
combined$last.name <- last.names

## pull 2nd part of each name after the comma and before space (title)
name.splits <- str_split(sapply(name.splits,"[", 2), " ")
titles <- sapply(name.splits,"[",1)
combined$title <- titles

unique(combined$title)

```

Clean up titles and group them together
```{r echo=FALSE}
titles[titles %in% c("Jonkheer.","Don.")] <- "Sir."
titles[titles %in% c("Lady.","Dona.","the")] <- "Lady."
titles[titles %in% c("Col.","Capt.","Major.")] <- "Officer."
titles[titles %in% c("Mme.")] <- "Mrs."
titles[titles %in% c("Mlle.","Ms.")] <- "Miss."

combined$title <- titles

```


Summarize them:

```{r echo=FALSE}
title.summary <- combined[1:891,] %>% 
  group_by(title, survived) %>% 
  summarize(count=n()) %>% 
  mutate(pct = count/sum(count))

c1 <- ggplot(title.summary, aes(x= title, y = pct, fill = survived)) +
  geom_bar(stat="identity", width = 0.7) +
  theme(legend.position="bottom") +
  coord_flip()


c2 <- ggplot(title.summary, aes(x= title, y = count, fill = survived)) +
  geom_bar(stat="identity", width = 0.7) +
  theme(legend.position="bottom") +
  coord_flip()

cowplot::plot_grid(c1,c2, labels = "AUTO", ncol=2)
```



#### sibsp

Predictive? **Yes**


```{r}
summary(combined$sibsp)
```

```{r echo=FALSE}
sibsp.summary <- combined[1:891,] %>% 
  group_by(sibsp, survived) %>% 
  summarize(count=n()) %>% 
  mutate(pct = count/sum(count))

d1 <- ggplot(sibsp.summary, aes(x= sibsp, y = pct, fill = survived)) +
  geom_bar(stat="identity", width = 0.7)

d2 <- ggplot(sibsp.summary, aes(x= sibsp, y = count, fill = survived)) +
  geom_bar(stat="identity", width = 0.7)

cowplot::plot_grid(d1,d2, ncol=1, labels = "AUTO")
```


#### parch

Predictive? **maybe**


```{r}
summary(combined$parch)
```

```{r echo=FALSE}
parch.summary <- combined[1:891,] %>% 
  group_by(parch, survived) %>% 
  summarize(count=n()) %>% 
  mutate(pct = count/sum(count))

d1 <- ggplot(parch.summary, aes(x= parch, y = pct, fill = survived)) +
  geom_bar(stat="identity", width = 0.7)

d2 <- ggplot(parch.summary, aes(x= parch, y = count, fill = survived)) +
  geom_bar(stat="identity", width = 0.7)

cowplot::plot_grid(d1,d2, ncol=1, labels = "AUTO")
```

#### sibsp + parch --> family.size

Predictive? **Maybe**

```{r}
combined$family.size <- combined$sibsp + combined$parch
family.size.summary <- combined[1:891,] %>% 
  group_by(family.size, survived) %>% 
  summarize(count=n()) %>% 
  mutate(pct = count/sum(count))

d1 <- ggplot(family.size.summary, aes(x= family.size, y = pct, fill = survived)) +
  geom_bar(stat="identity", width = 0.7)

d2 <- ggplot(family.size.summary, aes(x= family.size, y = count, fill = survived)) +
  geom_bar(stat="identity", width = 0.7)

cowplot::plot_grid(d1,d2, ncol=1, labels = "AUTO")
```



#### fare


```{r}
summary(combined$fare)
```

* Fare looks to be a sum of the fare paid for a **ticket**.
* Looks like there are instances where a ticket is shared by multiple passengers.  in these cases, the fare is the same across these passengers.
* This case is a wealthy family (mother and son) traveling with two non-family members (servants?).  Mother and son share 3 rooms together.


```{r}
combined[which(combined$ticket == "PC 17755"),c("passengerid", "age", "name", "fare", "ticket","cabin","family.size")]
```

* There are also some data suggesting that getting mroe rooms could be associated with higher fare


For a first stab, let's try normalizing for number of passengers by averaging the fare by number of people sharing that ticket.

```{r}
ticket.party.size <- rep(0, nrow(combined))
avg.fare <- rep(0.0, nrow(combined))
tickets <- unique(combined$ticket)

## for each ticket
for (i in 1:length(tickets)) {
  current.ticket <- tickets[i]
  ## get indexes associated with ticket
  party.indexes <- which(combined$ticket == current.ticket)
  ## calculate average fare per passenger for this ticket
  avg.current.fare <- combined[party.indexes[1], "fare"] / length(party.indexes)
  
  ## set avg fare and ticket party size for those people in that ticket party.
  for (j in 1:length(party.indexes)) {
    ticket.party.size[party.indexes[j]] <- length(party.indexes)
    avg.fare[party.indexes[j]] <- avg.current.fare
  }
}

combined$ticket.party.size <- ticket.party.size
combined$avg.fare <- avg.fare

```

we have 1 n/a for avg. fare, so will want to impute value in order to use in model.
```{r}
summary(combined$ticket.party.size)
summary(combined$avg.fare)
```

#### cabin


#### embarked


#### Summary of relevant features
* 


## (2) Exploratory modeling 1

## (3) Cross Validation

## (4) 



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
