---
title: "Example R Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## First time R, R Studio, and Woden Setup 

1. Install Latest versions of R & R Studio
* R: https://cran.r-project.org/mirrors.html
* R Studio: https://www.rstudio.com/products/rstudio/download/#download

2. First-time RJDBC setup for MacOs: After installing R, the steps on this page first need to be done to make the rJava and RJDBC packages work properly on MacOS

* http://www.owsiak.org/r-java-rjava-and-macos-adventures/

3. Then try installing the following packages in R, needed to interact with Woden (Amazon Redshift DB)

```{r}
## install packages needed to connect to redshift db
# install.packages("rJava")
# install.packages("RJDBC")
```


## Connect to Woden

### Load packages
```{r}
# library(rJava)
library(RJDBC)
```

### Open Connection to Woden

```{r echo=FALSE}
# update these
redshift_driver_path <- "/Users/jeffreyli/Downloads/RedshiftJDBC42-1.2.1.1001.jar"
woden.user <- "jeffli"
woden.pwd <- "hQtK5YuVqJ0cL9eu3Tx1I9KiCSjB8Gdf"
```

```{r}
# update these
redshift_driver_path <- "/Users/jeffreyli/Downloads/RedshiftJDBC42-1.2.1.1001.jar"
#woden.user <- "XX"
#woden.pwd <- "XX"

# connect to Amazon Redshift
driver <- JDBC("com.amazon.redshift.jdbc.Driver", redshift_driver_path, identifier.quote="`")
url <- paste("jdbc:postgresql://woden-redshift.p11a.com:5439/woden?user=",woden.user,"&password=",woden.pwd, sep="")

conn <- dbConnect(driver, url)
```

## Basic Analysis: Select, Create/Drop Statements in Woden

### Create temp table(s) using SQL statement(s), view its contents, and store it in an R data frame for analysis in R.
```{r}

# run drop / create temp table statements in redshift
dbSendUpdate(conn,
  "
  drop table if exists #test_jeff_table3;
  create table #test_jeff_table3 as 
  select case when user_computed_email_domain_group in ('GMAIL','YAHOO','AOL','MSN') then user_computed_email_domain_group else 'OTHER' end as domain_group
  , count(*)
  from dbm.jc_mailable_universe_matrix_classification
  group by 1
  order by 1;
  ")

# run select statement in redshift and view result
dbGetQuery(conn, 
  "
  select * from #test_jeff_table3;
  ")

# run query on temp table and store result in R dataframe
result <- dbGetQuery(conn, "select * from #test_jeff_table3")

```


Disconnect from Woden Connection
```{r results = FALSE}
# close connection
dbDisconnect(conn)
```